Methods for Large Scale Learning to Rank

### 2014-02-21 report/Chapters/Background.tex
referenties opgeschoond + definitie speed-up
begin uiteenzetting approaches L2R :)

%************************************************
\chapter{Introduction to Learning-to-Rank}
Different definitions of Learning-to-Rank exist. In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods. Figure \ref{fig:discriminative_training} describes the general process of machine learning. It shows training elements from an input space that are mapped to an output space using a model such that the difference between the actual labels of the training elements and the labels predicted with with the model are minimal in terms of a loss function.
\begin{figure}[!h]
\includegraphics[scale=0.26]{gfx/descriminative_training}
\caption{Machine learning framework for Learning-to-Rank, obtained from Liu\cite{Liu2007}}
\label{fig:discriminative_training}
\end{figure}\\
Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
\begin{description}
\item[Feature Based]{\emph{Feature based} means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning. That is, a Learning-to-Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}. \emph{Input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function} are hereby defined as follows:
	\begin{description}
	\item[Input Space]{contains the objects under investigation. Usually objects are represented by feature vectors, extracted according to different applications.}
	\item[Output Space]{contains the learning target with respect to the input objects.}
	\item[Hypothesis Space]{defines the class of functions mapping the input space to the output space. The functions operate on the feature vectors of the input object, and make predictions according to the format of the output space.}
	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
	\end{description}
	}
\end{description}

@@

The predictions and the loss function might either be defined for:
\begin{enumerate}
\item the relevance of a single document
\item the classification of the most relevant document out of a document-pair
\item the ranking of documents directly
\end{enumerate}
These three approaches are in literature respectively called the pointwise approach, the pairwise approach and the listwise approach. In the following sections we will further describe the three different approaches in Learning-to-Rank.

\section{Pointwise Approach}

\section{Pairwise Approach}
\section{Listwise Approach}
\chapter{Evaluation Methods}

\section{Normalized Discounted Cumulative Gain}
Cumulative gain, or its predecessor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
\subsection{Discounted Cumulative Gain}
The \ac{DCG}\cite{Burges2005} at a position $p$ is defined as\\
$DCG_p = \sum\nolimits_{i=1}^p \frac{2^{rel_i-1}}{log_2(i+1)}$\\
with $rel_i$ the graded relevance of the result at position $i$. The idea is that highly relevant documents that appear lower in a search result should be penalized (discounted). This discounting is done by reducing the graded relevance  logarithmically proportional to the position of the result.
\subsection{Normalized Discounted Cumulative Gain}
\ac{nDCG} normalizes the \ac{DCG} metric to a value in the [0,1] interval by dividing by the \ac{DCG} value of the optimal rank. This optimal rank is obtained by sorting documents on relevance for a given query. We can write the definition of \ac{nDCG} down mathematically as\\
$nDCG_p = \frac{DCG_p}{IDCG_p}$
\section{Expected Reciprocal Rank}
\ac{ERR}\cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$. \ac{ERR} is based on the reasoning that users are likely to stop exploring the result list once they have found a document that satisfied their information need. The \ac{ERR} metric is defined as the expected reciprocal length of time that the user will take to find a relevant document. \ac{ERR} is formally defined as\\
$ERR = \sum\nolimits_{r=1}^n \frac{1}{r} \prod\nolimits_{i=1}^{r-1}(1-R_i)R_i$\\
where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
An algorithm to compute \ac{ERR} is shown in Listing \ref{lst:err}. The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
\begin{lstlisting}[caption={Algorithm to compute the ERR metric, obtained from \cite{Chapelle2009}}, label={lst:err},language=Ada]
p <- 1, ERR <- 0
for r=1 to n do
	R <- R(g[r])
	ERR <- ERR + p * R/r
	p <- p * (1-R)
end for
return ERR
\end{lstlisting}
\section{Mean Average Precision}
\ac{AP}\cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
$AP = \frac{\sum\nolimits_{k=1}^{n}Precision(k)*relevance(k)}{\text{number of relevant docs}}$\\
With $k$ being the positions in the result set between 1 and $n$.
\ac{MAP} is the average \ac{AP} for a set of queries.\\
$MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$

>>>>

%************************************************
\chapter{A basic introduction to Learning-to-Rank}
Different definitions of Learning-to-Rank exist. In general, all ranking methods that use machine learning technologies to solve the problem of ranking are called Learning-to-Rank methods. Figure \ref{fig:discriminative_training} describes the general process of machine learning. It shows training elements from an input space that are mapped to an output space using a model such that the difference between the actual labels of the training elements and the labels predicted with with the model are minimal in terms of a loss function.
\begin{figure}[!h]
\includegraphics[scale=0.26]{gfx/descriminative_training}
\caption{Machine learning framework for Learning-to-Rank, obtained from Liu\cite{Liu2007}}
\label{fig:discriminative_training}
\end{figure}\\
Liu \cite{Liu2007} proposes a more narrow definition and only considers ranking methods to be a Learning-to-Rank method when it is \emph{feature based} and uses \emph{discriminative training}, which are itself defined as follows:
\begin{description}
\item[Feature Based]{\emph{Feature based} means that all documents under investigation are represented by feature vectors that reflect the relevance of the documents to the query.}
\item[Discriminative Training]{\emph{Discriminative training} means that the learning process can be well described by the four components of discriminative learning. That is, a Learning-to-Rank method has its own \emph{input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function}, like the machine learning process described by Figure \ref{fig:discriminative_training}. \emph{Input space}, \emph{output space}, \emph{hypothesis space}, and \emph{loss function} are hereby defined as follows:
	\begin{description}
	\item[Input Space]{contains the objects under investigation. Usually objects are represented by feature vectors, extracted according to different applications.}
	\item[Output Space]{contains the learning target with respect to the input objects.}
	\item[Hypothesis Space]{defines the class of functions mapping the input space to the output space. The functions operate on the feature vectors of the input object, and make predictions according to the format of the output space.}
	\item[Loss Function]{in order to learn the optimal hypothesis, a training set is usually used, which contains a number of objects and their ground truth labels, sampled from the product of the input and output spaces.}
	\end{description}
	}
\end{description}

@@

The predictions and the loss function might either be defined for:
\begin{enumerate}
\item the relevance of a single document
\item the classification of the most relevant document out of a document-pair
\item the ranking of documents directly
\end{enumerate}
These three approaches are in literature respectively called the pointwise approach, the pairwise approach and the listwise approach. These three approaches to Learning-to-Rank will be described in more detail further on in this part.

\chapter{Evaluating a ranking}
\section{Normalized Discounted Cumulative Gain}
Cumulative gain, or its predecessor discounted cumulative gain and normalized discounted cumulative gain, is one of the most widely used measures for effectiveness of ranking methods.
\subsection{Discounted Cumulative Gain}
The \ac{DCG}\cite{Burges2005} at a position $p$ is defined as\\
$DCG_p = \sum\nolimits_{i=1}^p \frac{2^{rel_i-1}}{log_2(i+1)}$\\
with $rel_i$ the graded relevance of the result at position $i$. The idea is that highly relevant documents that appear lower in a search result should be penalized (discounted). This discounting is done by reducing the graded relevance  logarithmically proportional to the position of the result.
\subsection{Normalized Discounted Cumulative Gain}
\ac{nDCG} normalizes the \ac{DCG} metric to a value in the [0,1] interval by dividing by the \ac{DCG} value of the optimal rank. This optimal rank is obtained by sorting documents on relevance for a given query. We can write the definition of \ac{nDCG} down mathematically as\\
$nDCG_p = \frac{DCG_p}{IDCG_p}$
\section{Expected Reciprocal Rank}
\ac{ERR}\cite{Chapelle2009} was designed based on the observation that \ac{nDCG} is based on the false assumption that the usefulness of a document at rank $i$ is independent of the usefulness of the documents at rank less than $i$. \ac{ERR} is based on the reasoning that users are likely to stop exploring the result list once they have found a document that satisfied their information need. The \ac{ERR} metric is defined as the expected reciprocal length of time that the user will take to find a relevant document. \ac{ERR} is formally defined as\\
$ERR = \sum\nolimits_{r=1}^n \frac{1}{r} \prod\nolimits_{i=1}^{r-1}(1-R_i)R_i$\\
where the product sequence part of the formula represents the chance that the user will stop at position $r$.\\
An algorithm to compute \ac{ERR} is shown in Listing \ref{lst:err}. The algorithm requires relevance grades $g_i$, $1 \le i \le n$ and mapping function $R$ that maps relevance grades to probability of relevance.
\begin{lstlisting}[caption={Algorithm to compute the ERR metric, obtained from \cite{Chapelle2009}}, label={lst:err},language=Ada]
p <- 1, ERR <- 0
for r=1 to n do
	R <- R(g[r])
	ERR <- ERR + p * R/r
	p <- p * (1-R)
end for
return ERR
\end{lstlisting}
\section{Mean Average Precision}
\ac{AP}\cite{Zhu2004} is an often used binary relevance judgement based metric that can be seen as a trade-off between precision and recall that is defined as\\
$AP = \frac{\sum\nolimits_{k=1}^{n}Precision(k)*relevance(k)}{\text{number of relevant docs}}$\\
With $k$ being the positions in the result set between 1 and $n$.
\ac{MAP} is the average \ac{AP} for a set of queries.\\
$MAP = \frac{\sum\nolimits_{q=1}^{Q}AP(q)}{Q}$

\chapter{Approaches to Learning-to-Rank}
%TODO: Pointwise, pairwise & listwise beter uitwerken
\section{Pointwise Approach}
Learn from each document in isolation.
Through:
	\begin{enumerate}
	\item regression, when relevance is real-valued
	\item classification, when relevance is categorical
	\item ordinal regression
	\end{enumerate}
\section{Pairwise Approach}
Pointwise approach Learning-to-Rank methods optimise real valued expected relevance, while evaluation metrics like \ac{nDCG} and \ac{ERR} are only impacted by a change in expected relevance when that change impacts a pairwise preference. The pairwise approach solves this drawback of the pointwise approach by regarding ranking as pairwise classification. Aggregating a set of predicted pairwise preferences into the corresponding optimal rank is shown to be a NP-Hard problem. 
\section{Listwise Approach}
Listwise ranking optimises the actual evaluation metric. The learner learns to predict an actual ranking itself without using an intermediate step like in pointwise or pairwise Learning-to-Rank. The main challenge in this approach is that most evaluation metrics are not differentiable. \ac{MAP}, \ac{ERR} and \ac{nDCG} are all discontinuous and non-convex functions, what makes them very hard to optimize.

### 2014-02-21 report/Chapters/Introduction.tex
begin uiteenzetting approaches L2R :)

To implement the Learning-to-Rank algorithms I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight. It which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.
The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo! Learning to Rank Challenge}\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR\cite{Qin2010} dataset and the LETOR successor MSLR-WEB30k.

>>>>

To implement the Learning-to-Rank algorithms I will use cloud based MapReduce implementation from Microsoft was used that is called HDInsight. It which is based on the popular MapReduce open source implementation Hadoop\footnote{http://hadoop.apache.org/}.
The algorithms that we include in the measurements will be determined based on experimental results on the \emph{Yahoo! Learning to Rank Challenge}\cite{Chapelle2011a}, the Yandex Internet Mathematics competition\footnote{http://imat-relpred.yandex.ru/en/}, the LETOR\cite{Qin2010} dataset and the LETOR successors MSLR-WEB10k and MSLR-WEB30k.
